---
title: "WDS, HW 2"
author:
- Group Member Srivastava, Abhay 
- Group Member Hong, Nathan
- Group Member Kang, Christiana
- Group Member O, Ella
- Group Member Wang, Felix
- Group Member Subramanian, Ramya
date: 'Due: 10:00PM, Friday, July 14, 2021'
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "hide", fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output
# check if you have ISLR package, if not, install it
if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(ISLR, readxl, tidyverse, magrittr, dplyr, ggplot2)
```


\pagebreak


# Objectives 

*The main purpose for this homework is to tell the difference between probability and statistics.* Using the Roulette game as illustration, we go through the main elements of probability and statstics. We then learn how to estimate unknown parameters in a population with precision, i.e., a confidence interval. Simulations help us to understand the interpretation of a confidence interval. 

- Continue to get used to `RStudio` and `RMarkdown`
- Probability
  - Random variables and their distributions
  - Population parameters: expected values and variances
  - Normal variables
  - Central limit theorem (CLT)

- Statistics
  - Sample means and sample variances
  - Use sample mean to estimate the population mean
  - Confidence intervals/hypothesis tests
  
- Packages/functions
    - `dplyr`
    - `ggplot`
    - `pnorm()`, `qnorm()`, `dnorm()`


# Review materials


- Study Module 2: Probability_Statistics_101


# Question 1: Probability

The following questions refer to the Roulette game. After the ball is spun it will land in one of the slots.

a.	What is the probability that the ball will land in any one slot?

1/38

b.	A player will bet \$1 each time. Let $X$ be the payoff.

    (i)	If the player bets on a single number, write down the distribution of $X$, i.e. the possible values of $X$ and the corresponding probabilities. Calculate the expected value of winning and the standard deviation.
  
 
    Possible Values of X = 35, -1
    
    Chances = 1/38, 37/38
    
    Expected Value=35 * 1/38 + (-1) * 37/38 = -.0526
    
    Standard Deviation
      -Variance=(35 + .0526)^2 * 1/38 + (-1 + .0526)^2 * 37/38 = 33.208  
      -sd = sqrt(33.208) = 5.76**

   
    (ii)	Repeat number (i) if the player bets on two different numbers.

    Possible Values of X = 34, -2
    
    Chances = 1/19, 18/19
    
    Expected Value=1/19 * 34 + (-2) * 18/19 = -.105
    
    Standard Deviation
      Variance=(34-(-0.105)^2) * 1/19 + (-1-(-0.105)^2) * 18/19 = 64.621
      sd = sqrt(64.621) = 8.04
 
 
    (iii)	If one plays the game by betting on a single number 1000 times, how much do you expect the person to win?
    
    Expected gain = -.0526 * 1000 = -52.6
    
    (iv)	Explain now why the casinos always make money on the roulette game.
    
    Because the expected value of winning is negative (-.0526).
    
c. Let's "play" the single number game using code. 

    (i) We bet 30 times each round. Put 30 payoffs into a vector and store into a variable. Use `set.seed(1)`.
    ```{r, results=T}
    set.seed(1)
    x <- rbinom(30, 1, 1/38)
    x <- x*36 - 1
    print(x)
    ```
    (ii) Plot the histogram of the 30 payoffs. What is the total payoffs for this round? 
    ```{r,results=T}
    hist(x)
    sum<-sum(x)
    print(paste("total payoffs:", sum))
    ```
    (iii) Let's play another 100 rounds. Each round we bet 30 times. Plot the histogram of total payoffs. What is the average payoff from these 100 rounds? Is this as expected? Use `set.seed(1)`.
    
```{r, results=T}
set.seed(1)
x <- rbinom(100, 30, 1/38)
x <- x*35 - 30 + x
hist(x)
x <- x/30
print(paste("average payoffs:", mean(x)))
#This is expected because it is .05 off from the true expected value.
```

# Question 2: Statistics 
    
This exercise is designed to help you understand the notion of confidence interval and understand confidence interval through simulations.

Let $Y$ be the average winning amount among 100 games of betting \$1 each time in the  Red-Black Roulette game. We know that $Y$ follows approximately a normal distribution with mean $\mu = -.0526$ and standard deviation $\sigma = .1$. We will start to record the game payoff each time but pretend that we do not know the true mean is -.0526. 

a. Generate 50 $Y's$ from $N(\mu = -.0526, \sigma =.1)$ with `set.seed(1)`. Construct a $95\%$ confidence interval for the unknown mean. State the confidence interval. Is the true value -.0526 inside the confidence interval?

```{r, results= T}
set.seed(1)
n_sim<-1 #number of simulations
mu_hat <- numeric(n_sim) # estimate
upper_ci <- numeric(n_sim) # lower bound
lower_ci <- numeric(n_sim) # upper bound
# Carry out the simulation
for (i in 1:n_sim){
y <- rnorm(50, mean= -.0526, sd = .1)
mu_hat <- mean(y)
mu_sd <- sd(y)
upper_ci[i] <- mu_hat + 1.96 * mu_sd/sqrt(50)
lower_ci[i] <- mu_hat - 1.96 * mu_sd/sqrt(50)
}
ci_df <- data.frame(
id = 1:n_sim,
lower_ci = lower_ci,
upper_ci = upper_ci)

print(ci_df)

mu<--.0526
if(between(mu,lower_ci,upper_ci) ){
  print("The true value -.0526 is inside the confidence interval.")
}
```


b. Repeat question 2a. 100 times to generate 100 samples (each has sample size 50 and follows $N(\mu = -.0526, \sigma =.1)$) and produce 100 95\% confidence intervals. How many of the confidence intervals capture the true mean -.0526? Display your 100 confidence intervals in a meaningful way. (Hints: using `geom_errobar()`)

```{r,echo=T,results=T}
set.seed(1)
n_sim<-100 #number of simulations
mu_hat <- numeric(n_sim) # estimate
upper_ci <- numeric(n_sim) # lower bound
lower_ci <- numeric(n_sim) # upper bound
# Carry out the simulation
for (i in 1:n_sim){
y <- rnorm(50, mean= -.0526, sd = .1)
mu_hat <- mean(y)
mu_sd <- sd(y)
upper_ci[i] <- mu_hat + 1.96 * mu_sd/sqrt(50)
lower_ci[i] <- mu_hat - 1.96 * mu_sd/sqrt(50)
}
ci_df <- data.frame(
id = 1:n_sim,
lower_ci = lower_ci,
upper_ci = upper_ci)

print(ci_df)

mu<--.0526
c <- 0
for (i in 1:length(lower_ci)){
  if(mu < upper_ci[i] & mu > lower_ci[i]){
    c = c+1
  }
}

p <- ggplot(data=ci_df, aes(x=id)) +
  geom_errorbar(aes(ymin=lower_ci, ymax=upper_ci))
p
```
c. Summarize the interpretation of a 95\% confidence interval based on the simulation in question 2b.  

If samples are repeatedly taken of a population, 95% of the confidence intervals of the samples will contain the population's true mean.

# Appendix: simulation code for confidence interval

To help out we provide the following R-chunk that provides 10 confidence intervals from 10 random samples. Modify it for your purpose accordingly.

```{r, eval = F, echo = T}
n_sim <- 10 # number of simulations
mu_hat <- numeric(n_sim)   # estimate
upper_ci <- numeric(n_sim)  # lower bound
lower_ci <- numeric(n_sim)  # upper bound

# Carry out the simulation
for (i in 1:n_sim){
  y <- rnorm(50, mean= -.0526, sd = .1)
  mu_hat <- mean(y)
  mu_sd <- sd(y)
  upper_ci[i] <- mu_hat + 1.96 * mu_sd/sqrt(50)
  lower_ci[i] <- mu_hat - 1.96 * mu_sd/sqrt(50)
}

ci_df <- data.frame(
  id = 1:n_sim,
  lower_ci = lower_ci, 
  upper_ci = upper_ci)
```

